### 2014-04-09 ApTranscriptome_TR.Rmd
Working on manuscript.
Updating analysis script to perform gene set enrichment analysis for

1. Intersection of genes by expression type from each colony
2. Set difference of genes by expression type from each colony

```{r temperature_response, eval=TRUE, echo=FALSE, results='asis'}
# table type of response
A22.type.table <- table(Ap.response.type[ , 'A22.type'])
Ar.type.table <- table(Ap.response.type[ , 'Ar.type'])
# overlap
overlap <- c(length(Ap.response.type[which(Ap.response.type$A22.type == "Bimodal" & Ap.response.type$Ar.type == "Bimodal"), "Transcript"]),
             length(Ap.response.type[which(Ap.response.type$A22.type == "High" & Ap.response.type$Ar.type == "High"), "Transcript"]),
             length(Ap.response.type[which(Ap.response.type$A22.type == "Intermediate" & Ap.response.type$Ar.type == "Intermediate"), "Transcript"]),
             length(Ap.response.type[which(Ap.response.type$A22.type == "Low" & Ap.response.type$Ar.type == "Low"), "Transcript"]),
             length(Ap.response.type[which(Ap.response.type$A22.type == "NotExp" & Ap.response.type$Ar.type == "NotExp"), "Transcript"]))

# make table
Ap.type.table <- rbind(A22.type.table, Ar.type.table, overlap)
rownames(Ap.type.table) <- c("A22", "Ar", "Both")

@@

pandoc.table(Ap.type.table, style="rmarkdown", caption = "Number of thermally-responsive transcripts with maximum expression at high, low, intermediate, both high and low (bimodal) temperatures or are not expressed for each colony and their overlap.")

@@

One of two tests, for *Ar* intermediate transcripts also finds greater constitutive expression in *Ar* than *A22*.

@@

@@

# perform enrichment analysis using parentchild method
Ap.BP.resultParentChild <- runTest(Ap.BP.GOdata, statistic = 'fisher', algorithm = 'parentchild')
Ap.BP.resultParentChild

# table results
Ap.BP.ResTable <- GenTable(Ap.BP.GOdata, parentchild = Ap.BP.resultParentChild, topNodes = 82)
#Ap.BP.ResTable
write.table(Ap.BP.ResTable, file = paste(resultsdir, "Ap_GO.BP_results.txt", sep=""), quote=FALSE, row.names=FALSE, sep = "\t")
pandoc.table(Ap.BP.ResTable)

@@

# Run GSEA
Ap.high.gsea <- gsea(genelist = Ap.geneList.high, geneID2GO = geneID2GO)
pandoc.table(Ap.high.gsea)
```

@@

# Run GSEA
Ap.low.gsea <- gsea(genelist = Ap.geneList.low, geneID2GO = geneID2GO)
pandoc.table(Ap.low.BP.ResTable)
```

@@

# Run GSEA
Ap.bim.gsea <- gsea(genelist = Ap.geneList.bim, geneID2GO = geneID2GO)
pandoc.table(Ap.bim.gsea)
```

@@

# Run GSEA
Ap.int.gsea <- gsea(genelist = Ap.geneList.int, geneID2GO = geneID2GO)
pandoc.table(Ap.int.gsea)
```

@@

# Run GSEA
A22.high.gsea <- gsea(genelist = A22.geneList.high, geneID2GO = geneID2GO)
pandoc.table(A22.high.gsea)
```

**Ar 'High' genes not in Ar**

@@

# Run GSEA
Ar.high.gsea <- gsea(genelist = Ar.geneList.high, geneID2GO = geneID2GO)
pandoc.table(Ar.high.gsea)
```

@@

# Run GSEA
A22.low.gsea <- gsea(genelist = A22.geneList.low, geneID2GO = geneID2GO)
pandoc.table(A22.low.gsea)
```

**Ar 'Low' genes not in Ar**

@@

# Run GSEA
Ar.low.gsea <- gsea(genelist = Ar.geneList.low, geneID2GO = geneID2GO)
pandoc.table(Ar.low.gsea)
```

@@

# Run GSEA
A22.bim.gsea <- gsea(genelist = A22.geneList.bim, geneID2GO = geneID2GO)
pandoc.table(A22.bim.gsea)
```

**Ar 'Bimodal' genes not in Ar**

@@

# Run GSEA
Ar.bim.gsea <- gsea(genelist = Ar.geneList.bim, geneID2GO = geneID2GO)
pandoc.table(Ar.bim.gsea)
```

@@

# Run GSEA
A22.int.gsea <- gsea(genelist = A22.geneList.int, geneID2GO = geneID2GO)
pandoc.table(A22.int.gsea)
```

**Ar 'Low' genes not in Ar**

@@

# Run GSEA
Ar.int.gsea <- gsea(genelist = Ar.geneList.int, geneID2GO = geneID2GO, plotpath = "results/Ar_int_BP_topGO_sig_nodes.pdf")
pandoc.table(Ar.int.gsea)
```

@@

>>>>

```{r temperature_response, eval=TRUE, echo=FALSE, results='asis'}
# table type of response
A22.type.table <- table(Ap.response.type[ , 'A22.type'])
Ar.type.table <- table(Ap.response.type[ , 'Ar.type'])
# overlap
overlap <- c(length(Ap.response.type[which(Ap.response.type$A22.type == "Bimodal" & Ap.response.type$Ar.type == "Bimodal"), "Transcript"]),
             length(Ap.response.type[which(Ap.response.type$A22.type == "High" & Ap.response.type$Ar.type == "High"), "Transcript"]),
             length(Ap.response.type[which(Ap.response.type$A22.type == "Intermediate" & Ap.response.type$Ar.type == "Intermediate"), "Transcript"]),
             length(Ap.response.type[which(Ap.response.type$A22.type == "Low" & Ap.response.type$Ar.type == "Low"), "Transcript"]),
             length(Ap.response.type[which(Ap.response.type$A22.type == "NotResp" & Ap.response.type$Ar.type == "NotResp"), "Transcript"]))

# make table
Ap.type.table <- rbind(A22.type.table, Ar.type.table, overlap)
rownames(Ap.type.table) <- c("ApVT", "ApNC", "Overlap")

@@

pandoc.table(Ap.type.table, style="rmarkdown", caption = "Number of transcripts with maximum expression at high, low, intermediate, both high and low (bimodal) temperatures or are not thermally-responsivefor each colony and their overlap.")

@@

The non-parametric test for both comparisions also finds greater constitutive expression in *Ar* than *A22*.

@@

```{r echo=FALSE, results='hide'}
# combine responsive.transcripts.ann with "type" information
Ap.response.type.sub <- 
responsive.transcripts.ann.type <- merge(responsive.transcripts.ann, Ap.response.type[,c("Transcript", "A22.type", "Ar.type")], by = "Transcript", all = TRUE)
str(responsive.transcripts.ann.type)
write.csv(responsive.transcripts.ann.type, file = "results/Ap_responsive_genes_20140409.csv", row.names = FALSE)
```

@@

# perform enrichment analysis using parentchild method
Ap.BP.resultParentChild <- runTest(Ap.BP.GOdata, statistic = 'fisher', algorithm = 'parentchild')
# Ap.BP.resultParentChild

# table results
Ap.BP.ResTable <- GenTable(Ap.BP.GOdata, parentchild = Ap.BP.resultParentChild, topNodes = 82)
dim(Ap.BP.ResTable)
#pandoc.table(Ap.BP.ResTable)

@@

# Run GSEA
Ap.high.gsea <- gsea(genelist = Ap.geneList.high, geneID2GO = geneID2GO, plotpath = "results/Ar_int_BP_topGO_sig_nodes.pdf")
#pandoc.table(Ap.high.gsea)
```

@@

# Run GSEA
Ap.low.gsea <- gsea(genelist = Ap.geneList.low, geneID2GO = geneID2GO)
#pandoc.table(Ap.low.BP.ResTable)
```

@@

# Run GSEA
Ap.bim.gsea <- gsea(genelist = Ap.geneList.bim, geneID2GO = geneID2GO)
#pandoc.table(Ap.bim.gsea)
```

@@

# Run GSEA
Ap.int.gsea <- gsea(genelist = Ap.geneList.int, geneID2GO = geneID2GO)
#pandoc.table(Ap.int.gsea)
```

```{r gsea_table, echo=FALSE, eval=TRUE, results='asis'}
## merge results into single table
Ap.high.gsea$Type <- "High"
Ap.low.gsea$Type <- "Low"
Ap.bim.gsea$Type <- "Bimodal"
Ap.int.gsea$Type <- "Intermediate"
# combine
Ap.gsea <- rbind(Ap.high.gsea, Ap.low.gsea, Ap.bim.gsea, Ap.int.gsea)
# reorder
Ap.gsea1 <- Ap.gsea[,c("Type", "GO.ID", "Term", "Annotated", "Significant", "Expected", "parentchild")]
colnames(Ap.gsea1)[7] <- "P"
# display
pandoc.table(Ap.gsea1, split.tables = "Inf")
```

@@

# Run GSEA
A22.high.gsea <- gsea(genelist = A22.geneList.high, geneID2GO = geneID2GO)
#pandoc.table(A22.high.gsea)
```

**Ar 'High' genes not in A22**

@@

# Run GSEA
Ar.high.gsea <- gsea(genelist = Ar.geneList.high, geneID2GO = geneID2GO)
```

@@

# Run GSEA
A22.low.gsea <- gsea(genelist = A22.geneList.low, geneID2GO = geneID2GO)
```

**Ar 'Low' genes not in A22**

@@

# Run GSEA
Ar.low.gsea <- gsea(genelist = Ar.geneList.low, geneID2GO = geneID2GO)
```

@@

# Run GSEA
A22.bim.gsea <- gsea(genelist = A22.geneList.bim, geneID2GO = geneID2GO)
```

**Ar 'Bimodal' genes not in A22**

@@

# Run GSEA
Ar.bim.gsea <- gsea(genelist = Ar.geneList.bim, geneID2GO = geneID2GO)
```

@@

# Run GSEA
A22.int.gsea <- gsea(genelist = A22.geneList.int, geneID2GO = geneID2GO)
```

**Ar 'Intermediate' genes not in A22**

@@

# Run GSEA
Ar.int.gsea <- gsea(genelist = Ar.geneList.int, geneID2GO = geneID2GO)
```

Make a table with results.

```{r gsea_table, echo=FALSE, eval=TRUE, results='asis'}
A22.high.gsea$Type <- "High"
A22.low.gsea$Type <- "Low"
A22.int.gsea$Type <- "Intermediate"
A22.bim.gsea$Type <- "Bimodal"
A22.gsea <- rbind(A22.high.gsea, A22.low.gsea, A22.int.gsea, A22.bim.gsea)
A22.gsea$Colony <- "ApVT"

Ar.high.gsea$Type <- "High"
Ar.low.gsea$Type <- "Low"
Ar.bim.gsea$Type <- "Bimodal"
Ar.int.gsea$Type <- "Intermediate"
Ar.gsea <- rbind(Ar.high.gsea, Ar.low.gsea, Ar.int.gsea, Ar.bim.gsea)
Ar.gsea$Colony <- "ApNC"

# combine
Ap.setdiff.gsea <- rbind(A22.gsea, Ar.gsea)
# reorder
Ap.setdiff.gsea <- Ap.setdiff.gsea[,c("Colony", "Type", "GO.ID", "Term", "Annotated", "Significant", "Expected", "parentchild")]
colnames(Ap.setdiff.gsea)[8] <- "P"

Ap.setdiff.gsea
```

@@

```{r rData, echo=FALSE}
# save RData for manuscript file
save(Ap.type.table, Ap.df, file = "ms.RData")
```
